---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# modeFinder

<!-- badges: start -->
<!-- badges: end -->

The goal of modeFinder is to compute the univariate mode estiamtes via simple empirical density estimates, kernel density estimates or Bernstein polynomials.

## Mode estimation via Bernstein polynomials

The R function of mode estimation via Bernstein polynomials is based on the draft of:

Liu, Bowen, and Sujit K. Ghosh. *"On empirical estimation of mode based on weakly dependent samples."* Computational Statistics & Data Analysis 152 (2020): 107046.

## Simple empirical mode estimation

It is natural to estimate the population distribution function by the empirical distribution function, $F_n$ given by

$$
F_n = \frac{1}{n} \sum^n_{i = 1} I \{X_i \leq x\}
$$

and for any $X^*$ such that $X^{(i)} < X^* < X^{(i + 1)}$, $i = 1, 2, \dots, n$, the slope of the empirical distribution function can be approximated by

$$
\frac{ F_n (X^{(i + 1)}) - F_n (X^{(i)})}{X^{(i + 1)} - X^{(i)}}.
$$

Consequently, we can give a simple mode estimator, $u^{(\hat{k})}$ based on the empirical distribution function by

$$
\hat{k} = \underset{0 \leq i \leq m^* - 1}{argmax} \frac{ F_n (u^{(i + 1)}) - F_n (u^{(i)})}{u^{(i + 1)} - u^{(i)}}.
$$

where $\{u^{(1)}, u^{(2)}, \dots, u^{(m^*)}\} = \{ \frac{1}{m^*}, \frac{2}{m^*}, \dots, \frac{m^*}{m^*}\}$. However, the simple empirical mode estimation does not collaborate with the possible smoothness characteristics of the underlying distribution function. 

## Installation

<!-- You can install the released version of modeFinder from [CRAN](https://CRAN.R-project.org) with: --> 

``` {r include = FALSE, echo = FALSE, eval = FALSE}
install.packages("modeFinder")
```

<!-- And -->  
You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("BowenNCSU/modeFinder")
```
## Example

This is a basic example which shows you how to solve a mode estimation problem:

```{r example}
library(modeFinder)
n = 1e6
set.seed(1)
data = rbeta(n, 2, 5)
## Return Bernstein polynomials based mode estimate
emp_mode(data)
## Return kenerl density mode estimate
emp_mode(data, smooth_option = "kernel")
## Return simple empirical mode estimate
emp_mode(data, smooth = FALSE)
```


### Note: 

macOS requires Xcode developer tools otherwise *xcrun: error* flags. Run the following: *xcode-select --install*.


